# textplex

<!-- badges: start -->
<!-- badges: end -->

The goal of textplex is to calculate text complexity using the algorithm by Tolochko & Boomgaarden (2019). Further validation of the algorithm is available in Tolochko, Song & Boomgaarden (2019). 

**Caveats:**

1. The current version uses udpipe to create dependency trees. Future version will provide an option to use [spacyr](https://github.com/quanteda/spacyr) (Benoit et al.) to maintain compatibility with the original implementation.
2. Super slow, but premature optimization is the root of all evil.
3. For algorithm by Benoit, Munger & Spirling (2019), use [sophistication](https://github.com/kbenoit/sophistication).

References:

1. Tolochko, P., & Boomgaarden, H. G. (2019). Determining Political Text Complexity: Conceptualizations, Measurements, and Application. International Journal of Communication, 13, 21.
2. Tolochko, P., Song, H., & Boomgaarden, H. (2019). “That Looks Hard!”: Effects of Objective and Perceived Textual Complexity on Factual and Structural Political Knowledge. Political Communication, 1-20.
3. Benoit, K., Munger, K., & Spirling, A. (2019). Measuring and explaining political sophistication through textual complexity. American Journal of Political Science, 63(2), 491-508.

## Installation

You can install the experimental version of textplex from github:

```{r eval=FALSE}
## install.packages("devtools")
devtools::install_github("chainsawriot/textplex")
```

## Examples

Calculating raw scores for English content.

```{r}
library(udpipe)
library(textplex)
## Download the english language model
## udpipe_download_model(language = "english")

## Load the english language model
udpipe_model <- udpipe_load_model("english-ewt-ud-2.4-190531.udpipe")

input_text <- c("A spectre is haunting Europe - the spectre of Communism.
All the powers of old Europe have entered into a holy alliance to exorcise
 this spectre; Pope and Czar, Metternich and Guizot, French Radicals
and German police-spies. Where is the party in opposition that has not been
 decried as communistic by its opponents in power? Where the Opposition that
 has not hurled back the branding reproach of Communism, against the more
 advanced opposition parties, as well as against its reactionary adversaries?",
"The greatest improvement in the productive powers of labour, and the greater
 part of the skill, dexterity, and judgment with which it is anywhere
 directed, or applied, seem to have been the effects of the division of
 labour. The effects of the division of labour, in the general business of
 society, will be more easily understood by considering in what manner it
 operates in some particular manufactures.")

calculate_textplex(input_text, udpipe_model)
```

Aber sprichst du Deutsch?

```{r}
## udpipe_download_model(language = "german")
de_udpipe_model <- udpipe_load_model("german-gsd-ud-2.4-190531.udpipe")

de_input_text <- c("Entschuldigung. Ich kann mit Ihnen auf Deutsch nicht
 sprechen, weil mein Deutsch sehr schlecht ist. Man sagt 'deutsche Sprache,
 schwere Sprache'. Ich glaube, dass ich nur Bahnhof verstehe.",
 "In mir drin ist alles rot, das Gegenteil von tot. Mein Herz es schlägt sich
 noch ganz gut. In mir drin ist alles rot und du bist ein Idiot, mein Freund.
 Du verschmähst mein süßes Blut.")

calculate_textplex(input_text, de_udpipe_model)
```

Fit the two-factor model.

```{r}
### it takes a long time. But you need to have a reasonable amount of data to fit the two-factor model.

library(dplyr)
library(psych)

## library(hcandersenr)
## hcandersen_en %>% group_by(book) %>% summarise(text = paste(text, collapse = " ")) %>% head(50) -> fairy

data(fairy)

fairy_rawscore <- calculate_textplex(fairy$text, udpipe_model)

fit <- fit_two_factor_model(fairy_rawscore)

fairy$syntactic_complexity <- fit$scores[,1]
fairy$semantic_complexity <- fit$scores[,2]

fairy %>% arrange(desc(syntactic_complexity)) %>% select(book, syntactic_complexity)
```

